---
title: "Calculate cost for meal items"
author: "Farid"
date: "19 August 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## AIM OF SHINY APPLICATION

- In this cafe, customer picks up items first and the receipt comprising charge, discount and final price is calculated at checkout counter using the shiny application
- Item code is input to the app for calculation of discount if eligible.
- As there can be a wide variation in costs for the same food type, only items below the reasonable cost for that food type is eligible for discounts. For example, Mee Siam comes in Economy ($3), Standard ($4) and Premium ($5). Only Economy and Standard is eligible for discount.

- Try out the app at <https://faridisme.shinyapps.io/Calculate-subsidy-for-selected-meal-items/>

## DESCRIPTION OF APP

- The app calculates the costs, discount and final price based on the eligible subsidy of the customer.

  1. The left-most panel lists the different menu items.
  2. Once menu items are selected, scan items by entering the item code into the scan forms in the middle panel.
    + Item codes within a scan form should be delimited by comma "," (or semicolon ";" or plus symbol "+");
    + For accurate subsidy calculation, scan set items (e.g. Nasi Lemak) based on meal sets, i.e. same nasi lemak set within same scan and different nasi lemak sets in different scans.
  3. Receipt and breakdown is updated in the right most column.


## SAMPLE RECEIPT OUTPUT

- For customer eligible for 50% discount initialised with the following items.
 
```{r, echo=TRUE, message=FALSE}
itemlist <- list("M1,L1,F1","N1,N3,N5") 
sublevel <- 50;

```


```{r, echo=FALSE, message=FALSE}
library(dplyr)

dfml <- read.csv("C:/Users/Farid/Dropbox/academic/Coursera/John Hopkins Specialisation in Data Science Course/Develop data products/Scan_ID/Masterlist.csv", stringsAsFactors = FALSE)
dffg <- read.csv("C:/Users/Farid/Dropbox/academic/Coursera/John Hopkins Specialisation in Data Science Course/Develop data products/Scan_ID/Foodgroupings.csv", stringsAsFactors = FALSE)

#Expand combinations or part there-of;
genNewRow <- function(x){
  y<-dim(x)[1]
  if(grepl(", ",x[y, c("Ingredient")])){
    #t <- list(); #initialist t as a list;
    t<- strsplit(x[y, c("Ingredient")],", ");
    t <- t[[1]][order(t[[1]])]# Re-arrange ingredients in alphabetical order;
    for (i in 1:length(t)){
      combi <- combn(t,i);
      for (j in 1:dim(combi)[2]){
        newcombi <- paste(combi[order(combi[,j]),j],collapse=", ")
        if(exists("savecombi")){
          savecombi <- c(savecombi, newcombi)
        } else { savecombi <- newcombi
        }
      }
    }
    newrow<- data.frame(Food=x[y,c("Food")],
                        Ingredient=savecombi,
                        Reasonable.Cost.cut.off=x[y,c("Reasonable.Cost.cut.off")]);
    rm(savecombi)
    return(newrow)
  } else {
    return(x)
  }
}
for (i in 1:nrow(dffg)){
  new_dffg_rows <- genNewRow(dffg[i,]);
  if(i!=1){
    dffgx <- rbind(dffgx, new_dffg_rows);
  } else {
    dffgx <- new_dffg_rows
  }
}

#Repeat for each itemlist;
for (i in 1:length(itemlist)){
  if(itemlist[[i]][1]=="Nil"|itemlist[[i]][1]==""){next}; #Skip if no inputs;
  a<-strsplit(itemlist[[i]][1], split="[,;+]+"); #Split based on , or ;
  df<-data.frame(Code=gsub(" ","",toupper(a[[1]]))); #Remove any white spaces;
  df$Code <- as.character(df$Code)
  df <- subset(df,df$Code!="Nil"); #remove 'Nil' items
  
  df$sn <- 1:nrow(df);
  df$Code <- as.character(df$Code);
  df <- merge(df, dfml, by="Code", all.x=TRUE) #merge with masterlist;
  df$Item[is.na(df$Item)] <- df$Code[is.na(df$Item)];
  df$Code<- NULL
  df$Ingredient[is.na(df$Ingredient)] <- "";
  df$Food[is.na(df$Food)] <- "Other item not on master list";
  df$Cost[is.na(df$Cost)] <- 0;
  
  #Separate items which are not ingredients;
  dfsave <- subset(df, df$Ingredient=="");
  df <- subset(df, df$Ingredient !="");
      
  if(nrow(df)>0){
    #aggregate data frame by combinations 
    df <- data.frame(df %>% 
                       summarise(sn=min(sn,na.rm=TRUE),
                                 Item=paste(Item[order(Ingredient)], collapse=", "),
                                 Ingredient=paste(unique(Ingredient[order(Ingredient)]), collapse=", "),
                                 Cost=sum(Cost, na.rm=TRUE)))
    }
  #merge with food groupings;
  df <- merge(df,dffgx, by="Ingredient", all.x=TRUE);
  df$Reasonable.Cost.cut.off[is.na(df$Reasonable.Cost.cut.off)] <- 0;
  df$Foodprevious <- NULL;
  
  #Rejoin with items which are not ingredients;
  dfsave$Ingredient <- NULL
  dfsave <- merge(dfsave,dffgx,by="Food",all.x=TRUE)
  dfsave$Ingredient[is.na(dfsave$Ingredient)] <- "";
  dfsave$Reasonable.Cost.cut.off[is.na(dfsave$Reasonable.Cost.cut.off)] <- 0;
  
  df <- rbind(dfsave, df);
  rm(dfsave)
  
  #Apply subsidy if under reasonable cost cut-off;
  df$Subsidy <- ifelse(df$Cost<=df$Reasonable.Cost.cut.off,sublevel/100*df$Cost,0)
  df$Price <- df$Cost - df$Subsidy
  
  #If multiple matches keep row with minimum price
  if(nrow(df)>0){
    df <- data.frame(df %>% group_by(sn,Item,Ingredient) %>%
                       slice(which.min(Price)))
  }
  
  df<-df[,c("sn","Item","Ingredient","Food","Cost","Subsidy","Price")]
      
  if(!exists("dftotal")){
    dftotal<- df;
    } else{
      dftotal<- rbind(dftotal,df)
    }
}
if(!exists("dftotal")){
      dftotal<-data.frame(sn=integer(),Item=character(),Food=character(),Cost=double(), Subsidy=double(), Price=double())
} else {
      dftotal$sn <- 1:nrow(dftotal); #Rename serial numbers in ascending order;
}

tab<- dftotal

str1 <- paste0("- Total charge is $",formatC(sum(tab$Cost, na.rm=TRUE), format="f",digits=2, big.mark=","))
str2 <- paste0("Customer is eligible for ",formatC(sublevel, format="f",digits=0, big.mark=","),"% discount");
str3 <- paste0("- Total discount is $",formatC(sum(tab$Subsidy, na.rm=TRUE), format="f",digits=2, big.mark=","));
str4 <- paste0("- Total price is $",formatC(sum(tab$Price, na.rm=TRUE), format="f",digits=2, big.mark=","));
str <- paste(str1, str2, str3, str4, sep="\n");
writeLines(str)

if(nrow(dftotal)>0){
  # Convert numbers to currency format;
  dftotal$Cost <- paste0("$",formatC(dftotal$Cost, big.mark=",",format='f', digits=2)); 
  dftotal$Subsidy <- paste0("$",formatC(dftotal$Subsidy, big.mark=",",format='f', digits=2));
  dftotal$Price <- paste0("$",formatC(dftotal$Price, big.mark=",",format='f', digits=2));
  }
names(dftotal)[names(dftotal)=="Subsidy"]<- "Discount"
```



```{r, echo = FALSE, fig.align='center', fig.cap='Receipt Breakdown'}
print.data.frame(dftotal[,c("Food","Cost","Discount","Price")], row.names=FALSE)

```

## UNDERLYING DATA

- *Masterlist.csv *
+ Master list of items with their costs is found in Masterlist.csv. The contents of the master list appears in the left-most panel of the app.

- *Foodgroupings.csv *
+ For each food, a reasonable cost cut-off has been determined to facilitate discount for eligible customers. When Food cost exceeds the reasonable cost cut-off, no subsidy is given.

- For more details on using R Markdown see <https://github.com/faridisme/Calculate-cost-and-discount-for-selected-meal-items>

